### 2020-11-07 paper.tex
coloque las observaciones de la Dra Vivechana y cambios bib

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx} % gráficos
\usepackage[noblocks]{authblk}
%\usepackage{authblk}
%\usepackage[•]{•}ckage{float}
%\usepackage{cite}% para contraer referencias
%\usepackage[usenames]{color}
%\usepackage{showkeys}
%\usepackage{ieeetr}
\usepackage{bm}
\usepackage{hyperref}
\usepackage[normalem]{ulem} %opcional para la correcciones
\newcommand{\abs}[1]{\lvert#1\rvert}
\usepackage{xcolor}
\newif{\ifinclude}
\newcommand{\notaL}[1]{{\color{blue}+L:#1+}}
\newcommand{\notaC}[1]{{\color{green}+C:#1+}}
\newcommand{\notaV}[1]{{\color{red}+V:#1+}}

\title{Model for Porous Silicon Structures with a Lateral Gradient}
\author[1]{C. A. Ospina de la Cruz}
\author[2]{V. Agarwal}
\author[3]{W. L. Mochán}

\affil[1]{Posgrado en Ingenier\'ia y Ciencias Aplicadas del Centro de
  Investigación en Ingenier\'ia y Ciencias Aplicadas (CIICAp-IICBA),
  Universidad Autónoma del Estado de Morelos (UAEM), Cuernavaca CP
  62209, México}
\affil[2]{Centro de Investigaci\'on en Ingenier\'ia y Ciencias
  Aplicadas,Universidad del Estado de Morelos, Av. Universidad 1001
  Col. Chamilpa, Cuernavaca, Morelos 62209, Mexico  }
\affil[3]{Instituto de Ciencias F\'isica ,Universidad Nacional
  Autonoma de Mexico, Av. Universidad S/N, Col. Chamilpa, 62210
  Cuernavaca, Morelos, Mexico}
\begin{document}

\maketitle
\notaV{Index Structures (NOTE: WE NEED TO CHANGE THE NAME;
  Probably "Analytical model for  porous silicon structures with 1D
  gradient refractive index)}
\notaL{The calculation is 2D, seems 1D due to the narrow geometry chosen.}
\begin{abstract}

A variety of optical devices and applications may be based on layered
systems which have a lateral gradient in both their refractive indices
and their thicknesses. They may be fabricated of porous silicon (PS) through an
electrochemical process in which the etching current depends on position. Thus, the possibility of
predicting the local current density $j$ and the corresponding
porosity $p$ and etching rate $v$ would be desirable for a systematic design of these
structures. In this work, we propose a
simple model for the calculation of $j$
as a function of position within an electrochemical cell. We prepared
some graded single layer PS samples, measured their optical properties through
ultraviolet-visible-near infrared reflectance
spectroscopy, \notaL{Should we mention microscopy?}, and related them
to their local porosity and thickness which, together with our model, were then used to obtain
the $p$ vs. $j$ and $v$ vs. $j$ calibration curves. 
\notaV{Thickness was additionally viewed through scanning electron microscopy}. We corroborated
our model by comparing the calibration curves for different samples. Thus, from
measurements on only one, or a few simple systems, we may obtain full
calibration curves which may be used to predict, design, and fabricate
more complex novel nonhomogenous devices with lateral gradients which can be used
for manifold applications.

Keywords: Porous Silicon; GRIN; Reflectance Spectra.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
Surfaces with a gradient in the refractive index (GRIN) are of
importance as they allow the engineering and tuning of the optical
phase along their surface, yielding applications such as flat
lenses. Furtheremore, there are important biological applications
which combine topographical features with index of refraction
gradients \cite{pierscionek2012gradient}. Preparation of structures
with gradients in other properties have also shown their
usefulness. \notaL{I believe the examples below are unrelated to GRIN,
  thats why I added the previous remark.} \notaV{Response : 
ALL the examples are related to GRIN} As an example, the
roughness dependent  response of cells has been tested on
materials used applications in medical implants
\cite{bai2020bioinspired}.
Gradients have also been employed for several
technological applications such as high-yield screening of catalysts,
sensing materials and biomolecules \cite{maier2007combinatorial, jayaraman2004construction,
potyrailo2008combinatorial}
Apart from some methods developed by S.E. Fosdick et al \cite{fosdick2010two},
%[JACS  2010, 132, 9226-9227],
synthesis of  Ag-Au alloy gradients on steel and chemical composition
gradients of CdS layers
on gold electrodes have been obtained made by Shannon’s
group\cite{wang2019fabrication}. \notaL{Is this Shannon a big shot, or
  why should we name him?} \notaV{R : Eliminated the name}
Among electrochemical methods, compositional and doping density changes
in  conducting polymers have been used to produce gradients using Indium Tin
Oxide electrodes \cite{inagi2010bipolar}
Electrochemically induced potential gradients have been shown to catalyze the
reactions and gradient doping of polymers \cite{qin2020bipolar, ishiguro2011gradient}.
As compared to the above mentioned techniques, a relatively economical,
fast and easy approach for the fabrication
of gradient materials has been the use  of asymmetrical
electrode configurations in the electrochemical synthesis of porous silicon \cite{sailor2012porous}.
This technique has the additional advantage of being compatible
with VLSI devices and integrable with microelectronics.

Although porous silicon (PS) initially stimulated the interest of the scientific
community due to its photoluminiscence in the visible range at room
temperature, it has presently
been recognized as a multifaceted optical material due to its large surface
area, biocompatibility, ease of fabrication and tunable refractive index.
Applications based on porous silicon now cover various fields such as
chemical sensors and biosensors \cite{kumar2020porous},
microelectronics and MEMS \cite{martinez2016dual},as well as a
range of optical \cite{estevez2014porous} and optoelectronic applications
\cite{ramadan2020fabrication}. Specifically,
the temporal variation of the current density results in a variation of
porosity along the depth, allowing the easy fabrication of different kinds
\cite{perez2018reflectivity} of 1D photonics crystals and other
structures. \notaL{Photonic crystals only in the periodic case.}
Although the conventional fabrication of PS yields laterally
homogeneous samples, with invariant structural and optical properties
along the surface, an asymmetric configuration \notaL{Is it clear what
  is asymmetric?} may produce porous
Silicon structures with a lateral gradient in properties such as
thickness, pore size and density, and thus, refractive index and
optical thickness. In this configuration, a small platinum electrode (cathode)
is positioned perpendicular  to the surface of a Si substrate (anode)
near one end \notaL{ At the center wouldn't be asymmetric: or the
  center} of the cell, resulting  in a current
density  which decreases as the distance to the electrode increases
 \cite{khung2008using, clements2011mesenchymal}.
The resulting porous surface can have pore sizes ranging from a few nanometers
to  few micrometers  \cite{wang2015screening}. The range of the dimensions of the pores and
the corresponding  thickness range on the same chip can be controlled by
adjusting the location of the electrode, the  anodizing current
and the etching time. Different
applications have been found for the resulting structures
\cite{collins2002determining}. For example, Sailor's group
\cite{collins2002determining, sailor2012porous} applied it
for the determination of protein size.
Additionally, gating of
the optical response through a variation in pH of the solution was shown to
trap/release biomolecules for possible drug delivery
applications, given the biocompatible nature of PS, . \notaL{How is this related to the gradient?}
\notaV{ The idea
here was to mention the usefulness of GRIN in portion 
size determination followed by its delivery through the variation of pH ; 
If you find this comment too way off, further applicability of protein 
size determination in drug delivery application can
 be removed "ADDITIONALLY.....applications" }
A similar asymmetric electrode configuration was used to fabricate
multilayered optical filters with lateral a gradient
\cite{li2005painting} and
for developing ethanol sensors \cite{sailor2012porous}.
Additionally it has been shown that the combination of thermal  oxidation  and
infiltration of TiO2  by \notaV{Atomic Layer Deposition} (ALD) \notaL{What is this acronym?} 
allows the manufacture
of transparent GRIN
optical elements with a high refractive index contrast \cite{ocier2017tunable}.
Kruger et al. \cite{krueger2018electrochemical} used
a porous silicon 1D GRIN structure infiltrated with polymer to
synthesize flat microlens arrays.
Recently, J Wang et al. has shown the fabrication of
a miniature spectrometer with PS based rugate filter using a
radial interfacial potential distribution \cite{wang2019fabrication}
On the other hand, the effect of different topographical
features present on the same chip, has also been
used to study the dependence of the adhesion of certain cells on the
surface topography \cite{khung2008using}.

For applications such as those mentioned above,
it would be very useful to be able to predict beforehand the spatially dependent
properties of the resulting structures. This would allow the design
and optimization of the desired devices.
In this work, we report a simple analytical model to quantitatively
analyze the spatial dependence of the current density during the
electrochemical on Si to produce PS structures. A calibration
procedure may then relate the current density to the porosity and the
etching rate. Thus, for a given etching current acting over a given
time, we may predict the position dependence of the index of
refraction and optical thickness of the resulting inhomogeneous
layer, thus allowing the prediction of the optical properties of
multilayered GRIN structure. The structure of the paper is the
following.
\notaL{Check when finished}. In Section \ref{sec:theory} we develop a
simple analytical model to calculate
the current density at any point over the GRIN structure under preparation.
In Section \ref{sec:procedures} we provide experimental details for the
synthesis of porous silicon samples with lateral refractive index gradient.
Section \ref{sec:discussion-results} includes experimental results and
it's comparison with numerically evaluated parameters. Finally, we
devote Section \ref{sec:conclusions} to our conclusions.

\section{Theory}
\label{sec:theory}
\label{sec:calc-curr-dens}
\notaL{check where acronyms are first defined}
A gradient in the optical properties of porous Silicon (PS) systems may be
obtained by electrochemically attacking a Si surface with a position dependent current
density ${\bm j(\bm r)}$, as the porosity and refractive index depend on
the current density, as well as the etching rate and the thickness of the
porous layer. Nevertheless, it is challeging to measure the
{\em local} value of $\bm j(\bm r)$ along the surface. This hampers the
design and optimization of well characterized reproducible GRIN PS
structures.\notaL{Tenemos referencias para afirmar esto?}
\notaV{NOT THAT I KNOW, so I changed the expression has been changed}
 Thus, we believe a
system for which the current density field may be
easily computed would be useful.
Though $\bm j(\bm r)$ may be numerically calculated for any
given arbitrary experimental setup,  here we propose a particularly
simple setup that allows an expression for the current in terms
of a series that is rapidly convergent and each of whose terms is a
simple analytic function. With this setup it is easy to characterize
the dependence of different properties of interest on the current, and
this allows the design and optimization of diverse GRIN PS devices.

We assume that our electrolytic cell has the shape of a rectangular
prism (see Fig. \ref{f:celda}) with a horizontal base of sides $a$ and $b$, and
filled with the electrolyte up to a height $c$. We assume that the
walls of the cell are insulating while the bottom is completely
covered with the sample, which is an
electrically grounded relatively good conductor. Within the electrolyte we position an electrode
which we assume is a thin conducting wire, insulated from the electrolyte
but for a small tip, which we approximate as a
point current source.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{celdan1.png}
  \caption{Prism shaped electrolytic cell with a rectangular base of
    sides $a$, $b$ filled with electrolyte up to a height $c$. The
    walls of the cell are insulating and the bottom is completely
    covered by the sample, which is assumed to be a good
    conductor. The tip of the electrode is a thin insulated wire
    with just the tip uncovered and located at $\bm r_0=(x_0,y_0,z_0)$ within the
    liquid. The current flows through the electrolyte from the sample
    to the tip.}
  \label{f:celda}
\end{figure}
Within the electrolyte the current density is
\begin{equation}
  \label{eq:j}
  \bm j=\sigma\bm E,
\end{equation}
with
$\sigma$ the conductivity, $\bm E=-\nabla\phi$ the electric field,
and $\phi$ the electric potential. In a
stationary situation
$\nabla\cdot \bm j = \sigma\nabla\cdot\bm E=0$. Thus, the potential
obeys Laplace's equation within the electrolyte,
\begin{equation}
  \label{eq:laplace}
\nabla^{2} \phi=0.
\end{equation}
Consider now an arbitrary closed surface
$\mathcal S$
within the electrolyte that surrounds completely the tip of the
electrode. Integrating $ \bm j$ over this surface we obtain
\begin{equation}
  \label{eq:I}
\int_{\mathcal S'} d\bm a \cdot \bm j = -I,
\end{equation}
where the prime on $\mathcal S'$ means we remove from the surface a
very small hole through which the wire that feeds the current to the
electrode gets
through, and the sign is consistent with the current direction in
Fig. \ref{f:celda}. Assuming the wire is narrower than any other relevant
distance in the system, we may interpret Eq. \eqref{eq:I} as an
integral over a closed surface of the current \eqref{eq:j} within the
electrolyte, ignoring the actual current within the wire. Thus
\begin{equation}
  \label{eq:gauss}
  \int_{\mathcal S} d\bm a \cdot \bm E = -\frac{I}{\sigma}
\end{equation}
Using Gauss's law, we interpret this equation as a source for the
potential in the form of a point charge
\begin{equation}
  q_0 = -\frac{I}{4 \pi \sigma}
\end{equation}
at the position $\bm r_0=(x_0, y_0, z_0)$ of the
tip of the electrode. Note that $q_0$ would be the total charge,
and it shouldn't be further screened through the permittivity of the
electrolyte.

The effective conductivity of the relatively thin sample in contact
with the grounded counter-electrode is large enough that we may assume its
surface $z=0$ is an equipotential.
On the other hand, no current can go across the insulating walls of
the cell, situated at $x=0$, $x=a$,, $y=0$ and $y=b$, nor through the
free surface of the liquid at $z=c$. Thus, the problem to solve is Poisson's
equation with a point charge source
\begin{equation}
  \label{eq:poisson}
  \nabla^2\phi=-4\pi q_0\delta^{(3)}(\bm r-\bm r_0),
\end{equation}
with mixed boundary conditions
\begin{equation}
  \label{eq:ground}
  \phi(x,y,0)=0,
\end{equation}
and
\begin{equation}
  \label{eq:E0}
E_x(0, y ,z)= E_x(a, y ,z)=E_y(x, 0 ,z)=E_y(x, b
,z)=E_z(x, y ,c)=0.
\end{equation}
Here, $\delta^{(D)}$ is Dirac's delta function in $D$ dimensions.
This problem may be solved readily using image charge theory \cite{jackson1975classical}.
The potential within the electrolytic cell coincides
within the region $0\le x\le a$, $0\le y \le b$, $0\le z\le c$
with that an infinite fictitious empty space but for the point charge
$q_0$ at the tip of the electrode at $\bm r_0$, and an array of its fictitious image
charges situated out of the cell. They include an image on the conducting bottom of the
cell, of the opposite charge $-q_0$ and situated at $(x_0, y_0,-z_0)$,
to guarantee that the bottom stays at potential zero.
There are further images on the four walls of the cell, of the same charge $q_0$
and situated at $(-x_0,y_0,z_0)$, $(x_0,-y_0, z_0)$, $(2a-x_0, y_0,
z_0)$ and $(x_0, 2b-y_0,z_0)$ so that, taken one at a time, no current goes across
the corresponding wall. Similarly, there is an image of the same charge $q_0$ at
$(x_0,y_0,2c-z_0)$ so that no current goes across the
surface of the electrolyte. Furthermore, each image charge has to be
further reflected by each of the aforementioned surfaces, yielding
infinitely many new images so that Eqs. \eqref{eq:ground} and
\eqref{eq:E0} hold when all charges are considered together.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{paper-1}
  \caption{Top view of the $z=z_0$ plane of the fictitious system formed by reflections of
    the electrolytic cell (light blue) showing the charge at the
    tip of the electrode (red dot) and its images on the walls of the
    rectangular cell, the images of its images and so on (blue
    dots). The walls are at $x=0,a$,
    $y=0,b$ (blue lines). The relation of some charges and their images is
    indicated by dashed arrows, with a bar to indicate the
    corresponding reflecting wall. The system may be interpreted as a
    periodic lattice with a rectangular unit cell of size $2a\times 2b$,
    one of which is indicated  by wide dashed lines, and with a basis of four equal
    charges $q_0$ at $(\pm x_0, \pm y_0, z_0)$.}
  \label{fig:topview}
\end{figure}
In Fig. \ref{fig:topview} we illustrate the real charge and all of
its images within the plane $z=z_0$. They form a periodic rectangular
lattice with a unit cell of size $2a\times 2b$ and with a basis of
four equal charges $q_0$ situated at the positions $\bm r_0$, $\bm
r_1=(-x_0, y_0, z_0)$, $\bm r_2=(x_0, -y_0, z_0)$ and $\bm r_3=(-x_0,
-y_0, z_0)$. In Fig. \ref{fig:sideview} we show a lateral view of the
image charges within the plane $y=y_0$.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{paper-2}
  \caption{Side view of the $y=y_0$ plane of the fictitious system formed by reflections of
    the electrolytic cell (light blue)
    showing the charge at the tip of the electrode (red dot), its
    images within the plane $z=z_0$ (blue dots), and its images at the
    surface of the sample and the top of the electrolyte (green
    dots). We use solid dots to denote images with the same charge $q_0$
    as that corresponding to the tip of the electrode, and open dots for those of
    opposite charge $-q_0$. The mirror planes that don't invert the charge
    are indicated by blue lines and the mirror plane that inverts the
    charge is indicated by a red line. The relation of some charges and their images is
    indicated by dashed arrows, with a bar to indicate the
    corresponding reflecting wall. The system may be interpreted as a periodic
    lattice with a rectangular unit cell of size $2a\times 4c$, one of
    which is indicated by wide dashed lines, and with a basis of four
    charges $q_0$ and another four charges $-q_0$. Within this cell,
    we indicate its $G=0$ contribution to the electric field (solid
    arrows, see text).}
  \label{fig:sideview}
\end{figure}
Each plane of charges upon reflection on the surface of the sample at
the bottom of the cell yields a
plane of image charges of the opposite sign, while each reflection
on the surface of the liquid yields a plane of charges of the same
sign. The side view may be interpreted as a periodic rectangular
lattice with a unit cell of size $2a\times 4c$ containing a basis of eight point
charges, four positive and four negative. A projection onto the $yz$
plane would similarly yield a rectangular lattice with a  $2b\times4c$
unit cell. Putting everything together, the problem is that of solving
Poisson's equation within an orthorhombic lattice of size
$2a\times2b\times4c$ with a basis of 16 point charges of charge $\pm
q_0$ arranged in positive and negative planes.

The electrostatic potential produced by an infinite array of
charges expressed as a sum in real space of Coulomb terms has ill
convergence properties. The potential can be obtained
using the Ewald summation technique
\cite{ewald1921calculation, belhadj1991molecular-dynamics,
  schmidt1997multipole, grzybowski2000ewald, toukmaji2000efficient,
  brodka2004ewald, campione2012ewald}, which yields
two rapidly converging series, one in real space and another one in reciprocal space.
Nevertheless, as we need the potential and the field to obtain the
current density on the surface of the sample on which  no
real nor image charges lie, we may perform a somewhat simpler planewise
summation, where we first obtain the potential and field produced by a
simple periodic lattice of charges using a fast convergent sum over the 2D
reciprocal vectors \cite{nijboer1957calculation, nijboer1958internal,
  de_wette1965internal, watson1981madelung, grindlay1981k}
\notaL{¿De qué se trata esta referencia que pusiste
  aquí: \cite{kratzer2019basics}?}, \notaC{Lo que quiero con este paper 
es resaltar los Conceptos básicos de la cristalografía, 
grupo de puntos cristalográficos, la red recíproca y algunos relaciones de la 
zona de Brillouin.}
then sum over the charges that form the basis
of the 2D {\em crystal} plane, and finally sum over all the
 planes.

We consider first a system made of identical {\em unit} charges
ocuppying the positions $\bm R$ of a 2D Bravais lattice $\{\bm R\}$,
which we will take as a rectangular lattice with lattice parameters
$2a$ and $2b$, and lying, for the time being, at the $z=0$ plane.
The potential $\phi$ it produces obeys
\begin{equation}
  \label{eq:poissonRed}
  \nabla^2\phi(\bm r)=-4\pi\sum_{\bm R}\delta^{(2)}(\bm r_\|-\bm
  R)\delta^{(1)}(z)=-\frac{4\pi}{A}\sum_{\bm G}e^{i\bm G\cdot\bm r_\|}\delta^{(1)}(z),
\end{equation}
where $\bm r_\|$ is the projection of the observation position $\bm
r=(x,y,z)$ onto the $xy$ plane, $\{\bm G\}$ is the
reciprocal lattice defined through $e^{i\bm G\cdot\bm R}=1$ for all
$\bm R$ and $\bm G$, $A=4ab$ is the area of the unit cell
and we used the Fourier representation of the 2D
periodically repeated delta function. We introduce a 2D Fourier representation for
the potential
\begin{equation}
  \label{eq:phifourier}
  \phi(\bm r)=\sum_{\bm G}\phi_{\bm G}(z)e^{i\bm G\cdot\bm r_\|}
\end{equation}
where $\phi_{\bm G}(z)$ is the Fourier coefficient of the potential at
the height $z$. Substitution in Eq. \eqref{eq:poissonRed} yields the
ordinary differential equation
 \begin{eqnarray}
 \frac{d^2}{dz^2}\phi_{\bm G}(z)-G^2\phi_G(z)=
   -\frac{4 \pi}{A} \delta^{(1)}(z),
 \end{eqnarray}
for each coefficient, an homogeneous equation for $z\ne0$ which may be trivially
solved. After applying boundary conditions at $z=0$ and regularity
conditions at infinity, we obtain
\begin{equation}
  \label{eq:phiG}
  \phi_{\bm G}(z)=
  \begin{cases}
    -\dfrac{2\pi}{A}\abs{z}&\text{if }G=0,\\
    \dfrac{2 \pi}{AG}  e^{- G \abs{z}}&\text{if }G\neq 0.
  \end{cases}
\end{equation}
The case $G=0$ corresponds to the potential produced by a uniformly charged plane,
while the case $G\ne 0$ is the potential produced by a sinusoidal
charge density on a plane, given by the
symmetric solution that decays exponentially as we get away, upwards
or downwards, from
the source plane. Substituting into Eq. \eqref{eq:phifourier} yields
\begin{eqnarray}
  \label{eq:phiG1}
 \phi(\bm r) = -\frac{2 \pi}{A}  \abs{z} +{\sum_{\bm G}}' \frac{2 \pi
   }{AG} e^{- G \abs{z}} e^{i \bm G\cdot\bm r_\|},
 \end{eqnarray}
where the prime indicates that the term $G=0$ should be ommited from
the sum.

We consider now the basis of our 2D lattice, with equal charges at the
four positions $(\pm x_0, \pm y_0,0)$. Each of the corresponding four
sublattices produces a potential as that in Eq. \eqref{eq:phiG1} but
shifted by the corresponding basis vector, yielding
\begin{equation}
  \label{eq:phiG4}
  \phi(\bm r) = -\frac{8\pi}{A} \abs{z}
    +{\sum_{\bm G}}' \frac{8\pi}{AG} e^{- G \abs{z}}\cos(G_x x_0)
    \cos(G_y y_0) e^{i \bm G\cdot\bm r_\|}.
\end{equation}
Finally, we shift the origin of Eq. \eqref{eq:phiG4} vertically to each of the planes shown in
Fig. \ref{fig:sideview} and multiply by the corresponding charge $q_0$
or $-q_0$ to obtain the contribution to the total potential. The charge $q_0$ corresponds to
the planes with heights of the form $z_0+4nc$ and $-z_0+(4n+2)c$, with
$n=\ldots-2,-1,0,1,2\ldots$ an arbitrary integer, while the charge
$-q_0$ corresponds to the planes with heights $z_0+(4n+2)c$ and
$-z_0+4(n+1)c$.

We concentrate our attention only on the region $0\le z<z_0$ above the
sample but below the electrode. From
Fig. \ref{fig:sideview} we see that the $G=0$
field has contributions only from the charges at $z_0$ and their
images at $-z_0$, and that this contribution is like that of a parallel plate
capacitor $E_{0z}=-4\pi q_0/ab=-16\pi q_0/A$, which corresponds to the
potential $\phi_0=16\pi q_0 z/A$. Other pairs of contiguous planes with opposite charges
contribute to the field elsewhere. The contributions to the potential
from the $G\ne0$ terms are simple geometric series that may be summed
over $n$ analytically and yield the total potential
\begin{equation}
  \label{eq:phitot}
  \begin{split}
    \phi(\bm r)=&\frac{16\pi q_0}{A}\biggl(z+{\sum_{\bm G}}'\frac{2\sinh
      Gc}{G \sinh 2Gc}\cosh G(c-z_0)\cos G_x x_0 \cos G_y y_0\\
    &\times \sinh Gz\, e^{i\bm G\cdot\bm r_\|}\biggr).
  \end{split}
\end{equation}

Notice that the terms of the sum above converge exponentially to zero
as $G$ increases for any $z$ such that $0\le z <z_0 \le c$, and thus
the sum is convergent. The divergence in the limit $ z
\longrightarrow z_0 $ corresponds to the singularity of the potential
at the position of a point charge, and it is not worrisome, as we are
interested in the field close to bottom $z=0$ of the cell.

The current density normal to the surface may now be obtained as
$j_\perp=-\sigma\partial\phi/\partial z$,
\begin{equation}
  \label{Eq:J}
  \begin{split}
    j_\perp(\bm r) =& \frac{I}{ab}\biggl(1+{\sum_{\bm G}}'\frac{2\sinh
      Gc}{\sinh 2Gc}\cosh G(c-z_0)\cos G_x x_0 \cos G_y y_0\\
    &\times \cosh Gz\, e^{i\bm G\cdot\bm r_\|}\biggr).
  \end{split}
\end{equation}
Eq. \eqref{Eq:J} is a very rapidly convergent expression that allows,
in particular, to calculate the current $j_\perp(x,y,0)$ that attacks
the substrate to produce the PS sample.
Integrating over the surface of the sample we verify
\begin{equation}
  \label{eq:intj}
  \int_0^adx\int_0^bdy\,j_\perp(\bm r)=I,
\end{equation}
as the contributions of opposite non-null reciprocal vectors cancel
out: as expected, all the charges that leave the sample find
their way  to the cathode across the electrolyte.

Using Eq. \eqref{Eq:J} we can correlate the local current density at
the sample as it is prepared with its geometric, structural and optical
properties.

\section{Procedures}
\label{sec:procedures}
\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{paper-3}\\[12pt]
  \includegraphics[width=.8\textwidth]{paper-4}\\[12pt]
  \includegraphics[width=.8\textwidth]{paper-5}

  \caption{Schematic experimental set-up for
    the fabrication of GRIN PS structures. We show a side view of the
    electrochemical cell with the shape of a
    rectangular prism (top panel). We indicate the Si wafer (dark
    gray), the teflon cell (light gray), the sealing o-ring (black
    circles), the insulated feeding cable (white lines within black
    lines), the uncovered platinum tip (white) within the
    electrolyte (blue), the electrical contact below the sample
    (yellow) and the controlled current source. We indicate some geometrical
    parameters of the arrangement $x_0$, $z_0$, $a$ and $c$
    and the current source and a few schematic current flux
    lines. Top and side schematic views of the resulting sample
    (middle and lower panels),
    illustrating eight regions $1\ldots8$ of decreasing porosity,
    illustrating the changes in pore radii,
    length, and density and the porosity gradient.}
  \label{fig:DE1}
\end{figure}

In Fig. \ref{fig:DE1} we show a schematic representation of our
experimental setup to manufacture GRIN PS structures.
As dicussed above, we fabricated an electrolytic cell with the shape of a
rectangular prism on the bottom of which we place a Si wafer which makes
electrical contact to a silver plate\notaC{ plata.}\notaL{Porqué
vivechna dice silver/copper? De qué era: latón, plata, cobre, una
aleación de cobre y plata? Era un alambre o una placa?}
\notaV{CRISTIAN :  POR FAVOR COLOCAS lo que usaste.}
which together make the anode. The
cathode consists of a platinum wire insulated except for a small region
at its tip, which we take as a point current drain.
We also show a schematic drawing of the expected GRIN PS structures,
with a gradient in the porosity, and in the thickness, density and depth
of the resulting pores.

\subsection{Fabrication of PS  GRIN Single Layers}
\label{sec:fabrication-ps-grin}
Single layer PS structures were fabricated through electrochemical
anodization of (100) oriented $p$-type crystalline Si wafer
(resistivity 0.002 - 0.005 $ \Omega\,\text{cm}$), under galvanostatic
conditions. The process was performed at room temperatured with an
electrolytic mixture of aqueous hydrofluoric
acid (HF) (concentration 48\% wt), glycerol (purity 99.8 \% wt) and
ethanol (purity 99.9\% wt) in a 3:7:1 volume ratio.
After the anodizing process, the samples were rinsed with
ethanol (purity: 99.9 $ \% $  wt). The electrolytic cell had the shape
of a rectangular prism as shown in Fig. \ref{fig:DE1}, with a base of
sides $a=2.01\,\text{cm}$ and $b=1.48\,\text{cm}$, with an uncovered sample area
of about $3\,\text{cm}^2$. The height of the liquid was set at
$c=1\,\text{cm}$. The cathode was a
platinum wire with a diameter of $0.41\,\text{mm}$ insulated with
teflon tape, with an uncovered
tip located at $\bm r_0=(x_0,
y_0,z_0)=(0.2\,\text{cm},0.74\,\text{cm},0.9\,\text{cm})$. \notaL{Checa todos estos datos}
We manufactured three samples $G_1$, $G_2$ and $G_3$ by applying a
current $I_1=5\,\text{mA}$, $I_2=10\,\text{mA}$ and
$I_3=20\,\text{mA}$, respectively, during a time of $t=250s$.

\subsection{Reflectance Measurement}
\label{sec:refl-meas}
We measured the reflectivity spectra at different positions over our
samples using a {\em Perkin Elmer Lambda 950}
UV-Vis-NIR spectrophotometer in its data collection absolute
reflectance mode, using a rectangular spot size of
$0.7\text{mm}\times2.0\text{mm}$, a slit width of $0.6\text{mm}$ and
a wavelength range from $1400\text{nm}$ to
$300\text{nm}$.\notaL{Vivechna: puedes checar si así se reportan estos parámetros?}
\notaV{CHANGED}
For interpretation of the data we used
the standard formula \cite{stenzel2015physics} for the reflectance of
a three media system:
an air environment  (0), the PS film (1) and the c-Si substrate (2),
\begin{equation}\label{Eq:ECMR}
  R=\left |\frac{r _ {_ {01}}r _ {_ {12}} e^{2i\psi}}{1+r _ {_
        {01}}r _ {_ {12}} e^{2i\psi}}\right |^2
\end{equation}
where $r_{01}$ and $r_{12}$ are the Fresnel reflectance
coefficients of the air/PS and PS/c-Si interfaces \cite{sattler2017silicon}
and $\psi=k_1^\perp d$, where $k_1^\perp$ is
the component of the wavevector perpendicular to the surface within
the porous layer of thickness $d$. The reflectance depends on the
thickness of the film and on the index of refraction $n_1=n_{\text{PS}}$ of the porous
layer, which we relate to the porosity through Bruggeman's
effective medium theory \cite{theiss1997optical},
\begin{equation}\label{Eq:Brugg}
 p\frac{1-\epsilon_1} {1+\epsilon_1} + (1-p)
  \frac{\epsilon_2 - \epsilon_1}
  {\epsilon_2+\epsilon_1}=0,
\end{equation}
where $\epsilon_2=\epsilon_{\text{Si}}$ is the dielectric function of
the c-Si substrate, and
$\epsilon_1=n_{\text{PS}}^2$ is the dielectric function of the PS layer.
We fitted the reflectance spectra using the porosity $p$ and the
thickness $d$ and a scale calibration factor $s$ as fitting
parameters, and thus, we obtained the porosity
$p$ and the etching rate $v=d/t$ for different values of the local
current density $j_\perp$ expected at several positions on different
samples.

\section{Discussion and Results}
\label{sec:discussion-results}
\subsection{Current Density}
\label{sec:current-density}
\begin{figure}
  \centering
  \includegraphics[width=1.2\textwidth]{j_Ivsxy.png}
  \includegraphics[width=\textwidth]{j_Ivsx.png}
  \caption{Normalized current density $j_\perp(x,y,0)/I$ calculated a
    the surface of the sample through
    Eq. (\ref{Eq:J}) for a
    system as in Figs. \ref{f:celda} and \ref{fig:DE1} for a cell of
    length $a = 2.01\,\text{cm}$ and width $b=1.48\,\text{cm}$, with
    the surface of the electrolyte at a heigth $c=1.0\,\text{cm}$, with a
    point-like electrode at $\bm r_0=(0.2\,\text{cm},
    0.74\,\text{cm},0.9\,\text{cm})$ (top panel). Normalized current
    along the center line $y=0.74\,\text{cm}$ (bottom
    panel).}
  \label{fig:DR1}
\end{figure}
In Fig. \ref{fig:DR1} we show the current density $j_\perp(x,y,0)$ calculated through
Eq. \eqref{Eq:J} at the surface of the sample for our cell, as shown
in Figs. \ref{f:celda} and \ref{fig:DE1} and described in
Sect. \ref{sec:fabrication-ps-grin}). For these parameters we obtained
a large range of values for $ j_\perp/I\sim 0.11-.72\,\text{cm}^{-2}$
that vary mostly along the $x$ direction. Of course, this would differ
for cells with different aspect ratios.

\subsection{Samples }
\label{sec:ps-grin-single}
\begin{figure}
 \centering
  \includegraphics[width=\textwidth]{G3I20proc}
  \caption{ Photograph of sample $G_3$ prepared as described in
    Sec. \ref{sec:fabrication-ps-grin} using the same parameters as
    in Fig. \ref{fig:DR1} with a current $ I_3=20 \text{mA}$,
    during a time $t=250\text{s}$. We indicate eight regions $x_i$
    $(i=1, 2,...8)$ centered at positions,
    $x'_1=0.2\text{cm}$,  $x'_2=0.4\text{cm}$, $x'_3=0.6\text{cm}$,
    $x'_4=0.8\text{cm}$, $x'_5=1\text{cm}$, $x'_6=1.2\text{cm}$,
    $x'_7=1.4\text{cm}$, $x'_8=1.6\text{cm}$, with $y_n=0.74\,\text{cm}$
    on which we measured reflectance spectra. Here,
    $x'_n=x_n+0.14\,\text{cm}$ is the position from the sealing o-ring.}
  \label{fig:SI3}
\end{figure}

In Fig. \ref{fig:SI3} we show a photograph of one of our samples,
$G_3$
prepared as described in Sec. \ref{sec:fabrication-ps-grin}, with the
same conditions as those corresponding to Fig.~\ref{fig:DR1} applying
a current
$I_3=20\,\text{mA}$ during a time $t=250\,s$. We notice a series of
visible interference fringes that get wider as we move towards the
right end of the sample, consistent with a inhomogeneous film that
gets thinner towards the right, and qualitatively consistent with
Fig. \ref{fig:DR1} that shows a current density that decays towards
the right. Equal color lines seem to correspond to iso-current lines of
Fig. \ref{fig:SI3}, except very close to the the walls of the cell,
where the film rapidly becomes very narrow. The reason is that there
is a small region below the cell walls where the electrolyte penetrates
up to the sealing o-ring (see upper panel of
Fig. \ref{fig:DE1}) and towards which the current leaks from the
neighborhood of the wall. The width of this region is
$w=0.14\,\text{cm}$. Thus, we may expect our formula \eqref{Eq:J}
to fail within that region and close to the border, at distances of the order of the radius
of the cross-section of the o-ring, of the order of a milimeter. The
figure shows eight positions
$x'_n=0.2\,\text{cm},0.4\,\text{cm},\ldots,1.6\,\text{cm}$ on which we
measured the near normal incidence reflection spectra of the
sample. Here, we defined $x'_n=x_n+w$ as the distance to the sealing o-ring,
where $x_n$ is the corresponding distance to the wall of the
cell.

From Eq. \eqref{Eq:J} we evaluate the
normalized current density $j_\perp/I$ expected at the positions $n$ indicated in
Fig. \ref{fig:SI3}.
(Table \ref{t:xj}).
\begin{table}
  \centering
  \begin{tabular}{rrr}
    \(x'_n\)(cm) & \(x\)(cm)& \(j_\perp/I\) (\(\text{cm}^{-2})\)\\
    from edge& from wall&\\
    \hline
    0.2 & 0.06 & 0.722\\
    0.4 & 0.26 & 0.672\\
    0.6 & 0.46 & 0.577\\
    0.8 & 0.66 & 0.464\\
    1.0 & 0.86 & 0.360\\
    1.2 & 1.06 & 0.275\\
    1.4 & 1.26 & 0.211\\
    1.6 & 1.46 & 0.167\\
  \end{tabular}
  \caption{Normalized current density $j_\perp/I$ expected at different positions $x'_n$.}
  \label{t:xj}
\end{table}

\subsection{Reflectance}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{R5vslambda}
  \caption{Reflectance spectra at different positions
    ($x'_n=0.2-1.6\,\text{cm}$, $y=0.74\,\text{cm}$)
    on sample $G_1$ ($I_1=5\,\text{mA}$). We show
  experimental results and results fitted through
  equations~\eqref{Eq:ECMR} and \eqref{Eq:Brugg}}.
  \label{fig:R5}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{R10vslambda}
  \caption{Reflectance spectra at different positions
    ($x'_n=0.2-1.6\,\text{cm}$, $y=0.74\,\text{cm}$)
    on sample $G_2$ ($I_1=10\,\text{mA}$). We show
  experimental results and results fitted through
  equations~\eqref{Eq:ECMR} and \eqref{Eq:Brugg}}.
  \label{fig:R10}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{R20vslambda}
  \caption{Reflectance spectra at different positions
    ($x'_n=0.2-1.6\,\text{cm}$, $y=0.74\,\text{cm}$)
    on sample $G_3$ ($I_3=20\,\text{mA}$) (Fig. \ref{fig:SI3}),
    We show
    experimental results and results fitted through
    equations~\eqref{Eq:ECMR} and \eqref{Eq:Brugg}}.
  \label{fig:R20}
\end{figure}

We measured the reflectance spectra  at the positions indicated in
Fig. \ref{fig:SI3} on our three samples $G_1$, $G_2$,
and $G_3$, prepared using the same procedure
(Sec. \ref{sec:fabrication-ps-grin}) but applying different currents
$I_1=5\,\text{mA}$, $I_2=10\,\text{mA}$, and $I_3=20\,\text{mA}$. The results are shown in
Figs. \ref{fig:R5}, \ref{fig:R10}, and \ref{fig:R20}.
We fitted the reflectance measurements using Eqs. \eqref{Eq:ECMR} and
\eqref{Eq:Brugg} using the local porosity $p$, and the thickness of the
film  $d$ as
adjustable parameters. We also fitted a scaling parameter $s$ to allow
for calibration variations among spectra. From the thickness we
obtained the etching rate $v=d/t$ at each position. Notice that the fits are reasonably
good. The contrast of the reflectance in the experiment is somewhat smaller
than in our fitted curves. We believe this is due to the finite width
$0.7\text{mm}\times2.0\text{mm}$ of the illuminating beam of our
spectrometer. Thus, as our GRIN sample is not macroscopically homogeneous and its
properties have a gradient, the
experimental results have contributions from regions with slightly
different porosities and thicknesses, averaging out the interference
maxima and minima. We expect microscopic inhomogeneities such as
interface roughness might also reduce the contrast through scattering
and its corresponding apparent dissipation \cite{missoni2020rough}.
The loss of contrast is most notable for the thicker regions
towards the left side of Fig. \ref{fig:SI3} and for smaller
wavelengths.

The results obtained from the reflectance fits are summarized in table
\ref{t:jpdv}. Notice that the scaling parameters are close to 1,
showing that the calibration was adequate. \notaL{Is this meaningful?}.

\begin{table}
\centering
\begin{tabular}{lrrrrrr}
  Sample/ & position & density & porosity &
  thickness & scale & rate\\
  $I$ (mA) &$x'_n$ (cm)&  $j$ ($\text{mA}/\text{cm}^2$) & $p$
       & $d$ ($\mu\text{m}$) & $s$ & $v$ (nm/s)\\
\hline
$G_1$/ 5 &0.2 & 3.61 & 0.494 & 0.340 & 0.849&  1.361 \\
         &0.4 & 3.36 & 0.571 & 0.390 & 0.979&  1.558  \\
         &0.6  & 2.88& 0.566 & 0.331 & 0.957&  1.325 \\
         &0.8 & 2.32 & 0.556 & 0.273 & 0.954&  1.093 \\
         &1.0 & 1.80 & 0.559 & 0.227 & 0.995&  0.894 \\
         &1.2 & 1.37 & 0.556 & 0.189 & 0.973&  0.754 \\
         &1.4 & 1.06 & 0.549 & 0.151 & 0.961&  0.605 \\
         &1.6 & 0.83 & 0.552 & 0.124 & 0.990&  0.495 \\
$G_2$/10 &0.2 & 7.22 & 0.591 & 0.635 & 0.856&  2.539 \\
         &0.4 & 6.72 & 0.644 & 0.693 & 0.991&  2.774 \\
         &0.6 & 5.77 & 0.638 & 0.591 & 0.952&  2.363 \\
         &0.8 & 4.64 & 0.626 & 0.533 & 0.980&  2.131 \\
         &1.0 & 3.60 & 0.599 & 0.459 & 0.939&  1.835 \\
         &1.2 & 2.75 & 0.590 & 0.404 & 0.945&  1.618 \\
         &1.4 & 2.11 & 0.575 & 0.338 & 0.936&  1.351 \\
         &1.6 & 1.67 & 0.571 & 0.296 & 0.948&  1.183 \\
$G_3$/20 &0.2 &14.45 & 0.686 & 1.204 & 1.149&  4.818 \\
         &0.4 & 13.44 & 0.713 & 1.193 & 1.078&  4.772 \\
         &0.6 & 11.54 & 0.703 & 1.044 & 1.059&  4.177 \\
         &0.8 &  9.29 & 0.626 & 0.752 & 1.006&  3.006 \\
         &1.0 &  7.20 & 0.625 & 0.538 & 1.010&  2.152  \\
         &1.2 &  5.50 & 0.593 & 0.439 & 0.962&  1.755 \\
         &1.4 &  4.22 & 0.584 & 0.363 & 0.962&  1.453 \\
         &1.6 &  3.33 & 0.591 & 0.339 & 1.011&  1.357 \\
\end{tabular}
\caption{\label{t:jpdv}
  Current density and fitted parameters for saveral positions on
  three samples:  porosity, thickness, scale factor and etching rate.}
\end{table}

From the data in Table \ref{t:jpdv} we can plot the fitted porosity as
a function of the calculated density current, as shown for our three
samples in Fig. \ref{fig:p}.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{pvsj}
  \caption{Porosity as a function of current density for three samples
    $G_1$, $G_2$, $G_3$. The data points for each sample are joined by
    thick lines. The arrows point to data that correspond to
    position $x'_1=0.2\,\text{cm}$, joined to the other data by thin
    lines (see text). We include linear fits to the data from
    sample $G_3$ (thick straight line) and to that from all the
    samples (thin straight line). }
  \label{fig:p}
\end{figure}
Notice that for each sample the data points lie around straight lines,
except for one outlier point in each corresponding to position
$x'_1$. As discussed in Sec. \ref{sec:ps-grin-single}, those points
were very close to a wall of the electrolytic cell, so they were
affected by leakage currents towards the underside of the wall.
For this reason, we exclude those points from the following
analysis. We made a linear fit to the data from sample $G_3$ and to
the data from the three samples. A comparison shows that a fit to the
data of just one sample, $G_3$, turns out to be a reasonable predictor for the
porosity of all the samples.  From that fit we obtain that for our
preparation conditions
\begin{equation}\label{eq:p}
  p\approx 0.013\frac{\text{cm}^2}{\text{mA}} j_\perp+0.53.
\end{equation}
The discrepancies may be due to the ill definition of the edge of the
sample (the position of the sealing o-ring) with respect to which we
measured the position of the illuminating spot. A rigid displacement along
$x$ yields mostly a horizontal shift of the results shown in
Fig. \ref{fig:p} and that can bring them into closer agreement.

Similarly, we can plot the fitted etching rate as
a function of the calculated density current, as shown for our three
samples in Fig. \ref{fig:v}.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{vvsj}
  \caption{Etching rate as a function of current density for three samples
    $G_1$, $G_2$, $G_3$. The data points for each sample are joined by
    thick lines. The arrows point to data that correspond to
    position $x'_1=0.2\,\text{cm}$, joined to the other data by thin
    lines (see text). We include linear fits to the data from
    sample $G_3$ (thick straight line) and to that from all the
    samples (thin straight line). }
  \label{fig:v}
\end{figure}
As for the porosity, there are outlier points corresponding to the
immediate proximity of the wall of the cell, but the remaining points
lie on straight lines for each sample, and they
may all be approximated by a linear fit to the data of just one
sample, $G_3$, from which we get
\begin{equation}\label{eq:v}
  v\approx 0.35\frac{\text{cm}^2}{\text{mA}}
  \frac{\text{nm}}{\text{s}}j_\perp + 0.092\frac{\text{nm}}{{s}}.
\end{equation}
The discrepancies, as in Fig. \ref{fig:p} might be due to our poor
definition of our reference position and may be eliminated by
introducing a rigid displacement along $x$ between the samples. For
multilayered samples, produced by modulating the current $I$ in time,
without changing wafers, these error source would be absent.

Using Eqs.  \eqref{Eq:ECMR}-\eqref{eq:v} we can calculate the
reflectance spectra for any
desired current density $j_\perp$ and etching time. This is
illustrated in Fig. \ref{fig:Rvsjl} for the parameters that correspond
to our samples.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{jlr}
  \caption{Reflection $R$ of a PS film as a function of the current density
    current $j_\perp$ and wavelength $\lambda$ for an etching time
    $t=250\,\text{s}$. Theory (left panel) and experiment (right
    panel, see text). We indicate with the numbers 1, 2, 3 the
    data corresponding to the corresponding samples $G_1$, $G_2$ and
    $G_3$.}
  \label{fig:Rvsjl}
\end{figure}
On the right hand side of the figure we display a quilt made by
patching together the experimental reflectance data corresponding to
Figs. \ref{fig:R5}- \ref{fig:R20}, ordered according to the
corresponding current density as presented in Table
\ref{t:jpdv}. Nevertheless, for the reasons mentioned above, we
eliminated the data corresponding to position $x'_1$ on all
samples. Furthermore, as the results corresponding to sample $G_2$ in
Figs. \ref{fig:p} and  \ref{fig:v} seem shifted with respect to those
of the other samples, in order to facilitate the visualization of the
results, we artificially shifted horizontally towards the
right their contributions to Fig. \ref{fig:Rvsjl} by
$1\,\text{mA}/\text{cm}^2$. Notice that the experimental results may
be grouped along bright and dark bands and that they correspond closely to
those expected theoretically, though the experimental bands are somewhat shifted,
especially for low current densities, and the minima are not as low.

The theoretical reflectance could as well be calculated using
\eqref{Eq:ECMR}-\eqref{eq:v} for more complicated structures, in which the
current $I$ is modulated in time, allowing the design and analysis of
non-trivial multilayered GRIN PS structures.

\section{Conclusions}
\label{sec:conclusions}
By choosing a simple geometry for our electrochemical cell, namely, a
rectangular prism, and employing a point-like electrode, we were able
to derive an expression for the calculation of the electric current
density along the surface of a Si sample. The expression was obtained
by using image charge theory and a planewise summation of Coulomb
potentials in 2D reciprocal space, and it is made up of
a rapidly converging sum of analytical terms, just a few of which
should be kept. This allows the characterization
of the porosity of a PS layer and the etching
rate for different values of the etching current density by making
optical measurements on several positions of just a
few samples. We illustrated the procedure with three samples on each of which we
measured reflectance spectra at eight different positions.
Through a optimization procedure we fitted the porosity and etching
rate corresponding to each current density.
We showed that an analysis of the data of just one of the samples, which we chose
as that with the largest current density range, yielded calibration
curves that allow the approximate prediction of the properties of the other
samples. We believe that the discrepancies we found between samples
are due to difficulties in defining a common reference position and
may be eliminated through a rigid shift along the $x$ direction.
The calibration curves can be used to design and calculate the optical
properties of other systems prepared under similar conditions, such as
multilayered GRIN systems from which photonic crystals, microcavities
and sensors with position dependent properties. This is the subject of
ongoing work. As the etching rate for the formation of porous silicon
and the resulting porosity are very sensitive to the growth conditions
and to the properties of the substrate, such as its doping,
there are no universal calibration curves. Thus, it is useful to have
a procedure that from just one sample can produce calibration curves
that may be employed for other samples grown in the same conditions.

\section*{Acknowledgments}
\label{sec:acknowledgments}
WLM acknowledges the support of DGAPA-UNAM under grant
IN111119. \notaL{Cristian debe agradecer a conacyt y faltan
  agradecimientos de Vivechna (quizás incluyendo al icf)}
\notaV{ acknowledges support from CONACyT Basic Sciences project A-S1 30393, 
PRODEP and ICF.}
\bibliographystyle{unsrt}
\bibliography{bibligrin} %Aquí ponen el nombre del archivo .bib
\end{document}